\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-14 trunk/arch.tex
Some updates and correction

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation \comment{Might be good to articulate what ``one'' step
    looks like}. The returned value indicates if a complete plan has
  been found. If so this part \comment{what does 'this part' mean? Can
    we sketch a figure or make it more formal?} of the plan is
  provided to {\em external} timelines of this reactor.

Both these deliberative functions have different scope and temporal
constraints. The \texttt{synchronize} call is considered as atomic and
will focus only on state inference for the current tick by identifying
the current value of {\em internal} state variables from the latest
updates on the {\em external} state variables. This process propagates
up in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process which is
focused on the future evolution of state variables; it can take
multiple steps spanning multiple ticks for a reactor to produce a plan
that is both complete and valid. When such plan is identified
\comment{how? Is there a terminating condition?} its {\em external}
timeline(s) are then dispatched to the corresponding reactors which
own these times and which in turn can start their own deliberation
\texttt{step}s to satisfy these new objectives. Consequently execution
occurs as these goals and plans propagate down in the reactor
hierarchy until they eventually become goals for the lowest level
reactor which -- instead of deliberating -- may just transform this
goal into a command for execution in hardware.

@@

While the two processes are separate, they share the same plan
internal to a reactor; consequently at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when the deliberation process eventually finds a
solution, synchronization is informed about generating the planned
state values \comment{should this read as ``...is informed about
  generated state values..'' instead?} on the {\em external} state
variables. This {\em external} plan defines the set of {\em goals} for
this state variable managed by another {\em reactor}.

@@

This assumption, while helpful to reduce the scope of the planning
problem, becomes problematic for planning within a highly dynamic
environment. In a typical robot this implies that the only point where
the planner can be informed about the current world state is prior to
the commencement of a planning cycle by giving to the planner the
initial state and current objectives. During the planning, it is
assumed that the agent can maintin the world as perceived by the
planner as stable. \cite{lemai04, lemai-chenevier2004} attempts to
reduce the impact of planning by allowing local plan repair in the
existing plan. In the eventuality that such repair is not feasible
they require that their robot comes to a full stop until (re-)planning
is complete. \comment{I may need more examples/refs here}. As Fig.
\ref{fig:tick-exec} shows, this assumption is problematic for our
architecture. Synchronization that tracks state evolution can occur
several times during the reactor deliberation process. By making
synchronization and deliberation share the same plan we provide a
solution that relaxes this static world assumption.

@@

For the sake of simplicity we introduce first how a reactor can track
and identify its state in the context where it does not have any
compelling need to deliberate. Assume here that we have a reactor that
have no future goal. {\em Need to develop the reason why it is a
  necessity: primarily to provide its internal state to whoever
  observe it but als simply to ensure that its current representation
  of the world is up to date and still consistent}

In the europa based reactor we introduce this requirement as a new
type of flaw for the europa framework. This flaw enforces at the
reactor to identify fully its internal state for the current tick. By
using this new flaw we can describe the synchronization process as
this general sequence :

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation}
  \caption{Illustration of the synchronization flaws in a reactor. The
    reactor receive new observations when they are produced by the
    owner(s) of its internal timelines. The line after the last token
    of each timeline represent the domain of possible values for the
    end of this token. At every tick $\tau$ the reactor needs to
    integrate the {\em External} state information it received and --
    based on its model $\mathcal{M}$ -- resolve its {\em Internal}
    state that will then be provided by the architecture to other
    reactors using these state variables.}
  \label{fig:synch:flaw}
\end{figure}

The resolution of internal state flaw can be resolved using one of the
following choices which are evaluated in the given order:

@@

\item Extend the previous state value to end after this tick (ie
  restrict its end time to $[\tau+1, \infty)$).

\item Start the next active token in the timeline and
  attempt to start it now (ie restrict its start time to the single
  value $\tau$).

>>>>

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of Europa for example pone step of
  deliberation correspod to resolve one sibngle flaw while ensuring
  that the resulting plan is not proven inconsistent. This implies
  that it also includes the backtracking of the search in such case. 
  The returned value by this step function is a boolean which
  indicates whether the reactor needs some extra step to deliberate or
  if this last step resulted on a complete plan for the reactor
  look-ahead. On the later, the part of the plan that concern this
  reactor {\em external} timelines is provided in order to be
  dispatched as goals to whichever reactor that is reponsible for
  these. \comment{By extending I may have saud things I developped
    below ... will have  to check}
\end{itemize}

Both these deliberative functions have different scope and temporal
constraints. The \texttt{synchronize} call is considered as atomic and
will focus only on state inference for the current tick by identifying
the current value of {\em internal} state variables from the latest
updates on the {\em external} state variables. This process propagates
up in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process which is
focused on the future evolution of state variables; it can take
multiple steps spanning multiple ticks for a reactor to produce a plan
that is both complete and valid. When such plan is identified
\comment{how? Is there a terminating condition? No more flaws but I
  washoping it would have been defined in europa section} its {\em external}
timeline(s) are then dispatched to the corresponding reactors which
own these times and which in turn can start their own deliberation
\texttt{step}s to satisfy these new objectives. Consequently execution
occurs as these goals and plans propagate down in the reactor
hierarchy until they eventually become goals for the lowest level
reactor which -- instead of deliberating -- may just transform this
goal into a command for execution in hardware.

@@

While the two processes are separate, they share the same plan
internal to a reactor; consequently at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when synchronization occurs after deliberation
steps, its starting point is the last partial plan resulting from
deliberation. This gives an initial frame to the problemm that helps
-- at least initial -- to target the synchronization problem by
ensuring that the new {\em external} state updates are compatible with
this plan. If so, the synchronization propagates these events in the
plan and can uses it as a help on identifying its {\em internal}
timelines based on this plan. If they are not compatibles, then
synchronization identifies a plan inconsistency  reflecting that the
current plan is not valid anymore. \comment{fpy: I am here and will
  develop further}

\comment{following is not what I intended to express whatsoever
  ... synchroniuzation od not deal with dispatching but execution
  tracking ... it is though helped by the structure of the exisiting
  plan and often identifies execution failures}
 Conversely, when the deliberation process eventually finds a
solution, synchronization is informed about generating the planned
state values \comment{should this read as ``...is informed about
  generated state values..'' instead?} on the {\em external} state
variables. This {\em external} plan defines the set of {\em goals} for
this state variable managed by another {\em reactor}.

@@

This assumption, while helpful to reduce the scope of the planning
problem, becomes problematic for planning within a highly dynamic
environment. In a typical robot this implies that the only point where
the planner can be informed about the current world state is prior to
the commencement of a planning cycle by giving to the planner the
initial state and current objectives. During the planning, it is
assumed that the agent can maintin the world as perceived by the
planner as stable. \cite{lemai04, lemai-chenevier2004} attempts to
reduce the impact of planning by allowing local plan repair in the
existing plan. In the eventuality that such repair is not feasible
they require that their robot comes to a full stop until (re-)planning
is complete. \comment{I may need more examples/refs here}. As Fig.
\ref{fig:tick-exec} shows, this assumption is problematic for our
architecture. Synchronization that tracks state evolution can occur
several times during the reactor deliberation process. By making
synchronization and deliberation share the same plan we provide a
solution that relaxes this static world assumption. \comment{Need to
  talk about IDEA and the fact that they had similar aspect even though
  they were never very clear nor well articulated in their papers}

@@

For the sake of simplicity we introduce first how a reactor can track
and identify its state in the context where it does not have any
compelling need to deliberate. Assume here that we have a reactor that
have no future goal. Even in this context, it is a core requirement in
T-REX (see \cite{py10}) to have each reactor to produce at every
single tick their {\em internal} state or at least, for the ones not owning
{\em internal} timelines, make sure that {\em external} observations
are taken into account for every tick. This requirement is enforced to
ensure that all the reactors of an agent share a consistent view of
the world as time advance.

Taking into account {\em external} state updates is manage d by
introducing these observation in the plan structure as factsand ensure
that they remain compatible sewith the current plan mainainted by the
reactor. The more delicate part of synchronization is to decide of the 
{\em internla} state values for the current tick. In the Shopping
example for example we have the following code snippet whihch helps
identify when we can \texttt{Buy} a product:
\begin{verbatim}
Agent::Buy {
  eq(duration, 10);
  // [...]
  contained_by(condition object.location.At currLocation);
  // [...]
}
\end{verbatim}
If \texttt{object.location} is {\em external} and the \texttt{Agent}
is {\em internal}. I may encounter the situation where I am at the
location where I can buy one (possibly among others) product. Already
there's here  a decision problem related to the state value of the
\texttt{Agent} timeline where I can either {\em Buy} a specific
product now, wait a little longer before doing so or just simply not
buy any product at this location.

In the europa based reactor we introduce this decision reuqirement 
as a new type of flaw for the europa framework. This flaw enforces at the
reactor to identify fully its internal state for the current tick. By
using this new flaw we can describe the synchronization process as
this general sequence :

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation}
  \caption{\small Illustration of the synchronization flaws in a reactor. The
    reactor receive new observations when they are produced by the
    owner(s) of its internal timelines. The line after the last token
    of each timeline represent the domain of possible values for the
    end of this token. At every tick $\tau$ the reactor needs to
    integrate the {\em External} state information it received and --
    based on its model $\mathcal{M}$ -- resolve its {\em Internal}
    state that will then be provided by the architecture to other
    reactors using these state variables.}
  \label{fig:synch:flaw}
\end{figure}

We first show how a reactor can track and identify its state in the
context where it does not have any compelling need to deliberate. We
assume that we have a reactor that has no future goal. \comment{Need
  to develop the reason why it is a necessity: primarily to provide
  its internal state to whoever observe it but also simply to ensure
  that its current representation of the world is up to date and still
  consistent}

We introduce this requirement as a new type of flaw (Section
\ref{sec:europa:arch}) for the \eu framework. This flaw forces the
reactor to identify its internal state for the current tick. By using
this flaw one can describe the synchronization process as:

@@

An internal state flaw can be resolved using one of the following
choices which are evaluated in order:

\item Extend the previous state value to end after this tick (\ie
  restrict its end time to $[\tau+1, \infty)$). \comment{first use of
    $\tau$, so needs to be defined earlier}

\item Start the next active token in the timeline 
  by restricting its start time to the single value $\tau$
